---
title: "Project1"
author: "Mubai Liu"
date: "9/24/2020"
output: html_document
---



# 1. Introduction
The *American National Election Studies* (ANES) are surveys of voters in the U.S. on the national scale. For each presidential election since 1948, ANES collects responses from respondents both before and after the election. The goal of ANES is to understand political behaviors using systematic surveys. ANES's data and results have been routinely used by news outlets, election campaigns and political researchers.


The *Time Series Cumulative Data* of ANES include answers, from respondents from different years, on selected questions that have been asked in three or more ANES' *Time Series* studies. Tremendous amount of efforts have been put into data consolidation as variables are often named differently in different years. 


```{r, include  = FALSE}
library(haven)
library(tidyverse)
anes_dat <-
    read_sav("../data/anes_timeseries_cdf.sav")
```



```{r process data, warning = FALSE, message = F, include = F}
Election_years = as.character((seq(1948, 2016, 4)))
anes_use=anes_dat%>%
  mutate(
    year=as_factor(VCF0004),
    turnout=as_factor(VCF0703),
    vote=as_factor(VCF0706),
    race=as_factor(VCF0105a),
    gender=as_factor(VCF0104),
    education = as_data_frame(VCF0140a),
    income = as_factor(VCF0114),
    region = as_factor(VCF0112)
  )%>%
  filter(year %in% Election_years) %>%
  select(year,
         turnout,
         vote,
         race,
         gender,
         education,
         income,
         region)

save(anes_use, file = "../output/data_story_use.RData")
```


